### 2025-01-18 src/main.tex

%%
%% This is file `sample-nime-paper.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,proceedings,bibtex,sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-nime-paper.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% History of NIME Template Updaters:
%% Modified by Florent Berthaut, Charles Martin, Yichen Wang, 21 November 2024
%% Modified by Adnan Marquez-Borbon 30 November 2022
%% Modified by Courtney Reed 28 November 2022
%% Modified by Joe Wright 14 December 2019
%% Modified by Niccolò Granieri 10 October 2018
%% Modified by Angelo Fraietta 23 December 2018
%% Modified by Angelo Fraietta 22 November 2018
%% Modified by Rodrigo Schramm on 22 September 2018
%% Modified by Luke Dahl on 17 October 2-17
%% Modified by Cumhur Erkut on <2016-10-11 Tue>
%% Modified by Edgar Berdahl on 5 November 2014
%% Modified by Baptiste Caramiaux on 25 November 2013
%% Modified by Kyogu Lee on 7 October 2012
%% Modified by Georg Essl on 7 November 2011
%%
%% And here's the template:
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{nimeart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{nimeart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf]{nimeart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{cc}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{}
%% These commands are for a PROCEEDINGS abstract or paper.
%% Make sure this is up to date with the correct edition of NIME.
\acmConference[NIME '25]{International Conference on New Interfaces for Musical Expression}{June 24--27,
  2025}{Canberra, Australia}
%% This suppresses the ACM Reference Format printing.
\settopmatter{printacmref=false}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmISBN{}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Augmentation of a Model Historical Harpsichord Mechanism for Haptic Enabled Interaction}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Matthew Hamilton}
% \authornote{Both authors contributed equally to this research.}
\email{matthew.hamilton2@unibo.it}
% \orcid{1234-5678-9012}
\author{Michele Ducceschi}
% \authornotemark[1]
\email{michele.ducceschi@unibo.it}
\affiliation{%
  \institution{Università di Bologna}
  \city{Bologna}  
  \country{Italy}
}

\author{Roberto Livi}
\author{Catalina Vincens}
\affiliation{%
  \institution{Museo San Colombano}
  \city{Bologna}
  \country{Italy}}

\author{Andrew McPherson}
\affiliation{%
 \institution{Imperial College}
 \city{London}
 \country{United Kingdom}}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
This work describes the design of an electronically augmented replica of a
16th-century harpsichord keyboard with a typical Renaissance Italian layout. The
keyboard is a collaboration between the authors and the Museo di San Colombano,
which houses the Tagliavini Collection of playable historical keyboard
instruments. The keyboard was commissioned for exhibition purposes and to
enhance the experiential authenticity of the collection. The keyboard emulates
the mechanics of a two-register harpsichord featuring a single manual. Unpitched
strings under tension replicate the mechanical resistance characteristic of the
original compass. An array of infrared LED/phototransistor pairs, connected to a
microcontroller, was installed within the interface to generate a continuous
stream of data corresponding to the jacks' motion. This paper details the design
principles, the development, open source implementation, and execution of
the installation for the San Colombano exhibition.
\end{abstract}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{interactive exhibit, augmented instrument, optical sensing, harpsichord}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{images/49-key-front.png}
  \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
  \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
  \label{fig:teaser}
\end{teaserfigure}

% \begin{teaserfigure}
%   \includegraphics[width=\textwidth]{images/side_w-o_sensors.png}
%   \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
%   \Description{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
%   \label{fig:teaser}
% \end{teaserfigure}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\include{article.tex}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
A project of this breadth developed across 3 institutes and 2 countries
cannot be carried out without the support and technical knowledge of many
skilled and passionate people.

We acknowledge the invaluable contributions of

\begin{itemize}
\item
  Università di Bologna
  \begin{itemize}
  \item
    Alma Labor
    \begin{itemize}
    \item
      Massimiliano Fraulini and the staff of Alma Labor at the Università
      di Bologna for their 3D printing expertise and generosity with
      electronics equipment
    \end{itemize}
  \item
    Fisica Tecnica
    \begin{itemize}
    \item
      Maurzio Chendi Fabrizio Casarini, Giulio Severo and Barbara
      Costantini
    \end{itemize}
  \end{itemize}
\item
  San Colombano
  \begin{itemize}
  \item
    Catalina Vincens
  \item
    Roberto Livi
  \end{itemize}
\item
  Edinburgh
  \begin{itemize}
  \item
    uCreate
    \begin{itemize}    
    \item
      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
      Edinburgh for their support with soldering facilities.
    \end{itemize}
  \item
    University of Edinburgh
    \begin{itemize}
    \item
      Richard Collins and the Digital Making Staff at the Edinburgh
      College of Art for 3D printing instruction and CNC support.
    \item
      Joe Hathaway of Digital Development at the College of Art for
      providing electronics components
    \item
      St Cecilia's Jenny Nex
    \end{itemize}
  \end{itemize}
\end{itemize}

\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

\end{document}
\endinput
%%
%% End of file `sample-nime-paper.tex'.